---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HAIgermany 🦠

[![Pkgdown](https://img.shields.io/badge/pkgdown-website-blue)](https://etc5523-2025.github.io/assignment-4-packages-and-shiny-apps-NathalieAaes/)

HAIgermany is an R package for exploring and visualizing healthcare-associated infections (HAIs) in Germany.
It provides tidy datasets, analysis functions, and an interactive Shiny app based on data from the ECDC Point Prevalence Survey (2011–2012).

---

## Installation

You can install the development version of HAIgermany from GitHub:

```{r}
# Install remotes if needed
install.packages("remotes")

# Install HAIgermany
remotes::install_github("ETC5523-2025/assignment-4-packages-and-shiny-apps-NathalieAaes/HAIgermany")
```

---

## Example Usage

### Load package and run the Shiny app:

```{r, eval=FALSE}
library(HAIgermany)

# Launch the interactive HAI Explorer app
run_HAIgermany_app()
```

### Explore datasets:

```{r, eval=FALSE}
# Distribution of McCabe scores
head(mccabe_scores_distr)

# Number of HAI patients by infection type and unit
head(num_hai_patients_tidy)
```

---

## Documentation

Full function and dataset documentation is available on the pkgdown site: [HAIgermany pkgdown site](https://etc5523-2025.github.io/assignment-4-packages-and-shiny-apps-NathalieAaes/)


